Prediction is Hard (Especially About the Future)

### 2015-08-08 Manuscript.tex
Updates from @rahlk's local machine.

\begin{figure}[t!]
	\small
	~\hrule~

	{\bf \fig{where}a: top down clustering:}

		The data is recursively divided in clusters using WHERE as follows:
		\begin{itemize}

@@

	{\bf \fig{where}b: finding distances:}

	In the \fig{where}a, to compute distances, we use
	the Euclidean measure recommended for
	case-based reasoning by Aha et al.~\cite{aha91};
	i.e. $\sqrt{\sum_iw_i(X_i-Y_i)^2}$ where $X_i,Y_i$
	where values are  normalized 0..1 for the range min..max and 
	$w_i$ is some importance weight given to attribute $i$.
	Usually, $w_i=1$ but it can be adjusted using, say,
	the feature weighting methods of \fig{where}d.

	{\bf \fig{where}c: 3 point geometry:}

	Let   $X,Y$ be two points joined by  a {\em Slope} of  length $c$.
	A third point, $Z$, has distances  $a=dist(Z,X)$ and
	$b=dist(Z,Y)$ to $X,Y$, respectively.
	According to the cosine rule,   $Z$ projects onto  $\overline{XY}$
	at distance $x=(a^2 + c^2 - b^2)/(2c)$ from $X$.
	Further, according to Pythagoras' theorem, $Z$ stands at a distance
	$y = \sqrt{a^2 - x^2}$ from the line $\overline{XY}$.

	{\bf \fig{where}d: feature weighting:}

	HERE's feature weighting algorithm
	comes from the CART regression tree learner~\cite{Breiman1984}.
	It sorts independent variables
	according to how well they can reduce the variability
	of a  numeric objective.
	If we split some column $i$ of independent numeric data  into $N=n_1 + n_2 + ..$ splits,
	then the expected
	value of the objective value of the split  is $w_i = \sum_i v_in_i/N$
	where $v$ is the standard deviation of the objective.
	A recursive procedure can  find those divisions $n_1,n_2,...$ by (a)~sorting a numeric column,
	then split at $\argmin_i w_i$, then recursing into each half.

	\caption{Sub-routines within HOW.}\label{fig:where}
\end{figure}

Decision tress categorically  sort the training data into a tree structure which can be viewed as model to reason about the data. In theory, this is a very desirable approach as it promises a

		{\bf Step1: Top down clustering using WHERE}

		The data is recursively divided in clusters using WHERE. For more information, see \fig{where}a.

		{\bf Step2: Distinguish between clusters using  decision trees}

@@

		{\bf Step3: Generating contrast sets from DT branches}
		\begin{itemize}
		\item Find the {\em current } cluster: take each test instance, run it down to a leaf in the DT tree.  
		\item Find the {\em desired} cluster: 
		\bi
		\item Starting at {\em current}, ascend the tree $lvl\in \{0,1,2...\}$ levels;
		\item Identify {\em sibling} clusters; i.e. leaf clusters that can be reached from level $lvl$ that are not {\em current }
		\item Using the {\em score} defined above, find the {\em better} siblings; i.e. those with a {\em score} less than $\epsilon=0.5$ times the mean score of {\em current}. If none found, then repeat for $lvl += 1$
		\item  Return the {\em closest} better sibling where distance is measured between the mean centroids of that sibling and {\em current}
		\ei
		\item Find the {\em delta}; i.e. the set difference between  conditions in the DT branch to {\em desired} and {\em current}. To find that delta:
		\bi
		\item
		For discrete attributes,  return the value from {\em desired}. 
		\item
		For  numerics, return the numeric difference. 
		\item
		For numerics  into ranges, return a random number selected from the low and high boundaries of the that range.
	\ei
	\ei
		~\hrule~
		\caption{CROSSTREES. Controlled by the parameters
		$\{\alpha, \beta, \gamma, \delta, \epsilon\}$ (set via engineering judgement).}
		\label{fig:xtrees_bare}
	\end{figure}

>>>>

%\begin{figure}[t!]
%	\small
%	~\hrule~
%	
%	{\bf \fig{where}a: top down clustering:}
%	
%		The data is recursively divided in clusters using WHERE as follows:
%		\begin{itemize}
%			
%			\item Find   two   distance cases,  $X,Y$
%			by picking any case $W$ at random, then setting $X$ to its most
%			distant case, then setting $Y$ to the case most distant from
%			$X$
%			(this requires only $O(2N)$ comparisons
%			of $N$ cases).
%			\item Project each case $Z$
%			onto a {\tt Slope} that  runs between $X,Y$ using the cosine
%			rule. 
%			\item Split the data at the median $X$ value of all cases and
%			recurses on each half  (stopping when
%			one half has less  than $\sqrt{N}$ of the original population).
%		\end{itemize}
%	
%	~\hrule~
%	
%	{\bf \fig{where}b: finding distances:}
%	
%	In the \fig{where}a, to compute distances, we use
%	the Euclidean measure recommended for
%	case-based reasoning by Aha et al.~\cite{aha91};
%	i.e. $\sqrt{\sum_iw_i(X_i-Y_i)^2}$ where $X_i,Y_i$
%	where values are  normalized 0..1 for the range min..max and 
%	$w_i$ is some importance weight given to attribute $i$.
%	Usually, $w_i=1$ but it can be adjusted using, say,
%	the feature weighting methods of \fig{where}d. 
%	
%	~\hrule~
%	
%	{\bf \fig{where}c: 3 point geometry:}
%	
%	Let   $X,Y$ be two points joined by  a {\em Slope} of  length $c$.
%	A third point, $Z$, has distances  $a=dist(Z,X)$ and
%	$b=dist(Z,Y)$ to $X,Y$, respectively.
%	According to the cosine rule,   $Z$ projects onto  $\overline{XY}$
%	at distance $x=(a^2 + c^2 - b^2)/(2c)$ from $X$.
%	Further, according to Pythagoras' theorem, $Z$ stands at a distance
%	$y = \sqrt{a^2 - x^2}$ from the line $\overline{XY}$. 
%	
%	~\hrule~
%	
%	{\bf \fig{where}d: feature weighting:}
%	
%	HERE's feature weighting algorithm
%	comes from the CART regression tree learner~\cite{Breiman1984}.
%	It sorts independent variables
%	according to how well they can reduce the variability
%	of a  numeric objective.
%	If we split some column $i$ of independent numeric data  into $N=n_1 + n_2 + ..$ splits,
%	then the expected
%	value of the objective value of the split  is $w_i = \sum_i v_in_i/N$
%	where $v$ is the standard deviation of the objective.
%	A recursive procedure can  find those divisions $n_1,n_2,...$ by (a)~sorting a numeric column,
%	then split at $\argmin_i w_i$, then recursing into each half.
%	
%	~\hrule~
%	
%	\caption{Sub-routines within HOW.}\label{fig:where}
%\end{figure}

		{\bf Step 1: Top down clustering using WHERE}
		The data is recursively divided in clusters using WHERE as follows:
		\begin{itemize}

@@

		{\bf Step 2: Distinguish between clusters using  decision trees}

@@

		{\bf Step 3: Generating contrast sets from DT branches}
		\begin{itemize}
		\item Find the {\em current } cluster: take each test instance, run it down to a leaf in the DT tree.  
		\item Find the {\em desired} cluster: 
		\bi
		\item Starting at {\em current}, ascend the tree $lvl\in \{0,1,2...\}$ levels;
		\item Identify {\em sibling} clusters; i.e. leaf clusters that can be reached from level $lvl$ that are not {\em current }
		\item Using the {\em score} defined above, find the {\em better} siblings; i.e. those with a {\em score} less than $\epsilon=0.5$ times the mean score of {\em current}. If none found, then repeat for $lvl += 1$
		\item  Return the {\em closest} better sibling where distance is measured between the mean centroids of that sibling and {\em current}
		\ei
		\item Find the {\em delta}; i.e. the set difference between  conditions in the DT branch to {\em desired} and {\em current}. To find that delta:
		\bi
		\item
		For discrete attributes,  return the value from {\em desired}. 
		\item
		For  numerics, return the numeric difference. 
		\item
		For numerics  into ranges, return a random number selected from the low and high boundaries of the that range.
	\ei
	\ei
		~\hrule~
		\caption{CROSSTREES. Controlled by the parameters
		$\{\alpha, \beta, \gamma, \delta, \epsilon\}$ (set via engineering judgement).}
		\label{fig:xtrees_bare}
	\end{figure}
